\bf Landslide: \\ Systematic Dynamic Race Detection in Kernel Space

### 2012-05-13 design.tex
add changebars

The happens-before relation expresses for each pair of transitions whether executing the first one is required to ``enable'' the second one. In order to establish the relation, Landslide uses the state of the scheduler runqueues (Section~\ref{sec:components-sched}), which are snapshotted at every decision point (Section~\ref{sec:components-explore}).

>>>>

\revision{The happens-before relation expresses for each pair of transitions whether executing the first one is required to ``enable'' the second one. In order to establish the relation,} Landslide uses the state of the scheduler runqueues (Section~\ref{sec:components-sched}), which are snapshotted at every decision point (Section~\ref{sec:components-explore}).

### 2012-05-13 evaluation.tex
add changebars

Of five student groups that we worked with during the user study, four groups completed the instrumentation process and got Landslide to explore at least a minimal decision tree. (A minimal decision tree is one resulting from decision points only on voluntary reschedules, which Landslide automatically identifies.) Of those groups, two found race conditions in their kernel. Additionally, Landslide helped all four groups find deterministic bugs - three groups had previously unknown problems with running the idle thread inappropriately, and the other group had a simple use-after-free.

>>>>

Of five student groups that we worked with during the user study, four groups completed the instrumentation process and got Landslide to explore at least a minimal decision tree. \revision{(A minimal decision tree is one resulting from decision points only on voluntary reschedules, which Landslide automatically identifies.)} Of those groups, two found race conditions in their kernel. Additionally, Landslide helped all four groups find deterministic bugs - three groups had previously unknown problems with running the idle thread inappropriately, and the other group had a simple use-after-free.

### 2012-05-13 futurework.tex
add changebars

We believe that a tool such as Landslide could change the way students learn such skills for the better, beyond simply increasing their likelihood of finding bugs during testing. The decision tree is a much more structured way of expressing a test case's concurrent behaviour than the way 15-410 currently teaches, which is ``think hard until you figure out the buggy interleaving''. Having students use Landslide, even to a small extent, might encourage them to think in this more structured way.\footnote{
In 15-410, before the kernel project, students also implement a user-space thread library. Just as Landslide could help the students during the kernel project, so too could a user-space systematic exploration tool help during the thread library project.}

>>>>

We believe that a tool such as Landslide could change the way students learn such skills for the better, beyond simply increasing their likelihood of finding bugs during testing. The decision tree is a much more structured way of expressing a test case's concurrent behaviour than the way 15-410 currently teaches, which is ``think hard until you figure out the buggy interleaving''. Having students use Landslide, even to a small extent, might encourage them to think in this more structured way.\footnote{
\revision{In 15-410, before the kernel project, students also implement a user-space thread library. Just as Landslide could help the students during the kernel project, so too could a user-space systematic exploration tool help during the thread library project.}}

### 2012-05-13 related.tex
add changebars

dBug\cite{dbug-ssv} is a systematic testing tool which makes use of DPOR. It focuses on multithreaded userspace applications, identifying when to preempt by interposing on \texttt{libc} and \texttt{pthreads} library calls. It uses a message-passing model for inter-thread communication and for establishing a happens-before relationship between transitions. The dBug project inspired the development of Landslide.

@@

RacePRO\cite{racepro} is a tool based in kernel space which uncovers ``process races'' (races involving interleaving system call patterns between multiple processes) using systematic exploration of multi-process interactions in user-space. They use system calls, such as filesystem accesses, as points to explore different interleavings of concurrent code, and detect bugs which can corrupt persistent system resources.

>>>>

dBug\cite{dbug-ssv} is a systematic testing tool which makes use of DPOR. It focuses on multithreaded userspace applications, identifying when to preempt by interposing on \revision{\texttt{libc} and \texttt{pthreads}} library calls. It uses a message-passing model for inter-thread communication and for establishing a happens-before relationship between transitions. The dBug project inspired the development of Landslide.

@@

RacePRO\cite{racepro} is a tool based in kernel space which uncovers ``process races'' \revision{(races involving interleaving system call patterns between multiple processes)} using systematic exploration of multi-process interactions in user-space. They use system calls, such as filesystem accesses, as points to explore different interleavings of concurrent code, and detect bugs which can corrupt persistent system resources.

### 2012-05-13 terminology.tex
add changebars

The following terms refer to parts of the guest kernel's execution.

@@

The following terms refer to parts of Landslide's analysis, regardless of the kernel under test's implementation.

>>>>

\revision{The following terms refer to parts of the guest kernel's execution.}

@@

\revision{The following terms refer to parts of Landslide's analysis, regardless of the kernel under test's implementation.}

### 2012-05-13 using.tex
add changebars

For deadlock detection, if the kernel has mutexes which loop around a call to \texttt{yield} (or similar), they must be annotated as described in section~\ref{sec:using-annotations}. If mutexes explicitly deschedule blocking threads, no special annotations are needed, because the other annotations do the job.

@@

The kernel should also use LMM (the \texttt{malloc} package in the 15-410 base code) for dynamic allocation, to avoid having to manually instrument a custom allocator.

@@

Some of these annotations may need to be used in ``sensitive'' regions of scheduler code, where it may not necessarily be clear at what point during the operation the annotation should go. Landslide assumes the scheduler already protects these regions by disabling preemption, and so the exact position of the annotation does not matter as long as it is within the preemption-disabled range.

@@

Both of these functions are predicates on the guest kernel's execution state (usually on the contents of its scheduler's data structures). Example implementations are given in Figure~\ref{fig:student-c}.

@@

\begin{itemize}
	\small
	\item \texttt{sched\_func} - When functions that make up the kernel's scheduler are identified, Landslide knows to ignore shared memory accesses from them. Accesses to scheduler data structures happen during {\em every} transition, so ignoring these is necessary for achieving any reduction at all.
	\item \texttt{ignore\_sym} - The user may wish to ignore memory accesses from other global data structures as well. For example, ignoring a mutex which is locked very frequently, but irrelevant to the actual test case, will result in an independence relation that better reflects what is actually being tested for.
	\item \texttt{within\_function} - If the user configures decision points to happen in common code paths, such as \texttt{mutex\_lock}, it will likely generate more branching than needed.\footnote{
For example, if testing \texttt{vanish}, there's no need to branch on a \texttt{mutex\_lock} in \texttt{exec}.}
This configuration option allows the user to whitelist functions so that Landslide will only decide if the current thread is ``within'' one of those functions.
	\item \texttt{without\_function} - Similar to above, but a blacklist. These two can be used together; later invocations take precedence over earlier ones.
\end{itemize}

>>>>

For deadlock detection, if the kernel has mutexes which loop around a call to \texttt{yield} (or similar), they must be annotated as described in section~\ref{sec:using-annotations}. \revision{If mutexes explicitly deschedule blocking threads, no special annotations are needed, because the other annotations do the job.}

@@

The kernel should also use LMM (the \texttt{malloc} package in the 15-410 base code) for dynamic allocation, \revision{to avoid having to manually instrument a custom allocator.}

@@

\revision{Some of these annotations may need to be used in ``sensitive'' regions of scheduler code, where it may not necessarily be clear at what point during the operation the annotation should go. Landslide assumes the scheduler already protects these regions by disabling preemption, and so the exact position of the annotation does not matter as long as it is within the preemption-disabled range.}

@@

\revision{Both of these functions are predicates on the guest kernel's execution state (usually on the contents of its scheduler's data structures). Example implementations are given in Figure~\ref{fig:student-c}.}

@@

\begin{itemize}
	\small
	\item \texttt{sched\_func} - When functions that make up the kernel's scheduler are identified, Landslide knows to ignore shared memory accesses from them. Accesses to scheduler data structures happen during {\em every} transition, so ignoring these is necessary for achieving any reduction at all.
	\item \texttt{ignore\_sym} - The user may wish to ignore memory accesses from other global data structures as well. For example, ignoring a mutex which is locked very frequently, but irrelevant to the actual test case, will result in an independence relation that \revision{enables much better state space reduction without losing the ability to find bugs in the parts of the kernel the user is interested in testing}.
	\item \texttt{within\_function} - If the user configures decision points to happen in common code paths, such as \texttt{mutex\_lock}, it will likely generate more branching than needed.\footnote{
For example, if testing \texttt{vanish}, there's no need to branch on a \texttt{mutex\_lock} in \texttt{exec}.}
This configuration option allows the user to whitelist functions so that Landslide will only decide if the current thread is ``within'' one of those functions.
	\item \texttt{without\_function} - Similar to above, but a blacklist. These two can be used together; later invocations take precedence over earlier ones.
\end{itemize}
